\mytitle

### 2021-12-14 paper.tex
improve table head
extend abstract
reduce strictness

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage[german]{authblk}
\usepackage[format=plain,labelfont={bf}]{caption}
\usepackage{listings}
\usepackage[shortcuts]{extdash} % \-/
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\setkomafont{disposition}{\bfseries}

@@

An algorithm for the computation of the real dilogarithm is presented,
which allows for the use of instruction\-/level parallelism.

@@

The development of modern CPUs with support for instruction\-/level
parallelism (ILP) opens the opportunity for a further reduction of the
runtime of the calculation of the dilogarithm.  To make use of ILP,
the algorithm to calculate the dilogarithm must be formulated
appropriately in order to allow for the parallel execution of multiple
floating\-/point operations.  In this paper an algorithm for the
calculation of the real dilogarithm $\Li$ is presented, which makes
use of ILP, resulting in a reduced runtime compared to the algorithms
presented so far.

@@

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures with
    different compilers.}
  \begin{tabular}{lcc}
    \toprule
    & i7-4700MQ & i7-5600U \\
    Implementation & \texttt{g++} 10.2.1 & \texttt{g++} 9.3.0 \\
    \midrule
    this paper & $0.031$ & $0.026$ \\
    Morris (DILOG 0011) \cite{morris} & $0.035$ & $0.026$ \\
    Cephes \cite{cephes} & $0.044$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.054$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.079$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.125$ & $0.071$ \\
    GSL \cite{gsl} & $0.330$ & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper is fastest
among the ones shown in the table.  The next-to fastest implementation
is provided by Cephes, which is only $\sim 20\%$ slower.  The short
runtime of Cephes stems from the fact that Cephes also uses a rational
function approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.

>>>>

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb}
\usepackage[german]{authblk}
\usepackage[format=plain,labelfont={bf}]{caption}
\usepackage{listings}
\usepackage[shortcuts]{extdash} % \-/
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{multirow}
\setkomafont{disposition}{\bfseries}

@@

An algorithm for the computation of the real dilogarithm is presented,
which allows for the use of instruction\-/level parallelism when
implemented on appropriate CPUs.

@@

The development of modern CPUs with support for instruction\-/level
parallelism (ILP) opens the opportunity for a further reduction of the
runtime of the calculation of the dilogarithm.  To make use of ILP,
the algorithm to calculate the dilogarithm must be formulated
appropriately in order to allow for the parallel execution of multiple
floating\-/point operations.  In this paper an algorithm for the
calculation of the real dilogarithm $\Li$ is presented, which makes
use of ILP, potentially resulting in a reduced runtime compared to
most algorithms presented so far.

@@

In \tabref{tab:runtime} the runtime of different C implementations is
compared on two different CPU architectures.  On each architecture the
implentations are compiled with \texttt{g++} with optimization
\texttt{-O2}.  The shown runtime denotes the total wall clock time for
the calculation of $\Li(x)$ for $10^6$ uniformly distributed random
values $x\in[-5,5]$.
%
\begin{table}[tb]
  \centering
  \caption{Runtime in seconds for the calculation of $\Li(x)$ for
    different implementations on different CPU architectures with
    different compilers.}
  \begin{tabular}{lcc}
    \toprule
    \multirow{2}{*}{Implementation} & i7-4700MQ & i7-5600U \\
    & \texttt{g++} 10.2.1 & \texttt{g++} 9.3.0 \\
    \midrule
    this paper & $0.031$ & $0.026$ \\
    Morris (DILOG 0011) \cite{morris} & $0.035$ & $0.026$ \\
    Cephes \cite{cephes} & $0.044$ & $0.034$ \\
    Algorithm 327 \cite{koelbigDilog} & $0.054$ & $0.035$ \\
    Chebyshev \cite{luke} & $0.079$ & $0.060$ \\
    Algorithm 490 \cite{ginsberg} & $0.125$ & $0.071$ \\
    GSL \cite{gsl} & $0.330$ & $0.155$ \\
    \bottomrule
  \end{tabular}
  \label{tab:runtime}
\end{table}
%
We find that the implementation presented in this paper is fastest
among the ones shown in the table.  The next-to fastest implementation
is provided by Cephes, which is only $\sim 20\%$ slower.  The short
runtime of Cephes stems from the fact that Cephes also uses a rational
function approximation, where the numerator and the denominator can be
evaluated in parallel.  However, Cephes evaluates the polynomials in
the numerator and denominator using Horner's scheme, which is purely
sequential.  Furthermore, for $x>1$ Cephes calculates $\ln(x)$ twice,
which leads to a further slowdown.
